EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-14 draft.tex
Comment about asynchronous implementations
@jj  please change at will
sentence clarified

% Probably it's not the time for this, but maybe it would be interesting to mention that asynchronous implementations in most cases need a change in the underlying algorithm, while synchronous might not need it - JJ
Moreover, parallel implementations can run in two different ways: synchronous and
asynchronous. When operations run in parallel but every node must maintain synchrony with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  master/worker model, the master needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other, so there's no time lost waiting for other populations to
arrive to the synchronization point; results arriving from other workers are incorporated as soon as they are received. Following the previous example, now the master operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

% I have moved these sentences here, since they seem to fit the best.
Beyond the choice of design methodology and implementation, researchers need to
consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
the interaction between them, the search area of each population, and the search
strategy and parametrization of each population. We must take into account
these high-level issues when designing a parallel architecture.

@@

Docker container images host the stateless functions responsible for running 
a search strategy on several populations, % clarify "on several populations" Sequentially? At the same time? - JJ
these functions could also be ported
directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.

>>>>

% Probably it's not the time for this, but maybe it would be interesting to 
% mention that asynchronous implementations in most cases need a change in 
% the underlying algorithm, while synchronous might not need it - JJ 
% TO DO - Mario 
Moreover, parallel implementations can run in two different ways: synchronous and
asynchronous. When operations run in parallel but every node must maintain synchrony with
the other operations,  all operations need to finish to continue to the next
step. For instance, in a  master/worker model, the master needs to wait for all
workers to finish their operations before moving to the next iteration.  In
contrast, in an asynchronous execution,  operations are not synchronized with
each other, so there's no time lost waiting for other populations to
arrive to the synchronization point; results arriving from other workers are 
incorporated as soon as they are received. Following the previous example, now the master operation can
continue to the next iteration, even if a worker has not finished its work.  In
the literature, we can find many asynchronous algorithms
\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
and scalability. In the particular case of asynchronous and cloud-native
multi-population algorithms, recent works propose an asynchronous communication
through a central repository \cite{sofea:cec2012, JSON} or message queues
\cite{salza2019speed, guervos2018introducing}, in this work we follow this
practice.

@@

% I have moved these sentences here, since they seem to fit the best.(OK - M)
Beyond the choice of design methodology and implementation, researchers need to
consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
the interaction between them, the search area of each population, and the search
strategy and parametrization of each population. We must take into account
these high-level issues when designing a parallel architecture.

@@

Docker container images host the stateless functions responsible for 
running a search strategy on populations they receive as messages. 
A single container can pull many messages on its lifetime, processing one at a time. 
% clarify "on several populations" Sequentially? At the same time? - JJ 
% done - M
These functions could also be ported
directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
% Need to improve the next paragraph or maybe move it to another place - Mario
Researchers then define the configuration of services used by
a multi-population algorithm using a YAML file. Any researcher can deploy and
start all the services from this configuration file with a single command. Once
the services are up, multiple instances of an experiment can be sent to a queue
for their execution. The platform includes containerized services that are
compatible with other container orchestration technologies like Kubernetes and
Docker Swarm. The services included are a Redis Server used for the Message
Queue implementation and logging and an Experiment Controller implemented in
Python. Developers of population processing functions have the freedom to
implement new population-based algorithms or operators in the language they
choose inside a container, as they only need to subscribe to a channel of the
Message Queue and have the ability to read the population and parameters as JSON  
file.
